# (PART) Model parameterization {.unnumbered}

# Climate inputs and weather generator

## Preperation and loading data

To add weather and climate data to our SWAT project, we will use `svatools`

### Required packages

```{r, loadpacks_climweath, message=FALSE}
require(euptf2) # devtools::install_github("tkdweber/euptf2") 
require(svatools) # devtools::install_github("biopsichas/svatools")
require(readr)
require(readxl)
require(sf)
require(mapview)
```

### Load in the file(s) and load svatools template

First we load in the template and fill it in with our values and rename it to "cs10_weather_data.xlsx". This is not done within R.

```{r}
#temp_path <- system.file("extdata", "weather_data.xlsx", package = "svatools")
# /// fill out this template and save "cs10_weather_data.xlsx" ///
```

We are using the following projection for this project:

```{r}
epgs_code = 25832
```

Now we can load it in with Svatools.

```{r}
met_lst <- load_template(template_path = "model_data/input/met/cs10_weather_data.xlsx", epgs_code)
```

### Proof the station

```{r, svatool_plot_weather}
plot_weather(met_lst, "PCP", "month", "sum")
```

Checking the location of the station:

```{r}
basin_path <- "model_data/input/shape/cs10_basin.shp"
basin <- st_transform(st_read(basin_path), epgs_code) %>%
  mutate(NAME = "Basin")

stations <- st_transform(met_lst$stations, epgs_code)
mapview(stations) + mapview(basin)
```

## Creating the weather generator

We only have one weather station, which means only one weather generator.

```{r}
wgn <- prepare_wgn(met_lst, 
                   TMP_MAX = met_lst$data$ID1$TMP_MAX, 
                   TMP_MIN = met_lst$data$ID1$TMP_MIN, 
                   PCP = met_lst$data$ID1$PCP, 
                   RELHUM = met_lst$data$ID1$RELHUM, 
                   WNDSPD = met_lst$data$ID1$WNDSPD, 
                   MAXHHR = met_lst$data$ID1$MAXHHR, 
                   SLR = met_lst$data$ID1$SLR)

write.csv(wgn$wgn_st, "model_data/input/met/as_wgn_st.csv", row.names = FALSE, quote = FALSE)

write.csv(wgn$wgn_data, "model_data/input/met/as_wgn_data.csv", row.names = FALSE, quote = FALSE)
```

# Adding weather data to SWAT project

Currently, this throws an error -- Cannot tell why, but might not be an issue because it still writes the files that we need. Because the buildR script is currently broken, we cannot evaluate the following code

```{r, eval=FALSE}
db_path <- "model_data/cs10_setup/optain-cs10/...sqlite"
add_weather(db_path, met_lst, wgn)
```

